Analysis paper

### 2022-05-12 analysis/analysis.Rmd
update manuscript with review suggestions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE)
options(knitr.kable.NA = '')
library(here)
library(magrittr)
library(scoringutils)
library(knitr)
library(kableExtra)
library(data.table)
library(ggplot2)
library(ggridges)
library(RColorBrewer)
library(patchwork)
library(stringr)
library(covid.german.forecasts)
library(grates)
library(ggdist)
library(ggthemes)
library(scales)
library(dplyr)
library(tidyr)
library(forcats)

@@

# helper function for labeling plots
scale_fn <- function(x) {
  ifelse(x >= 5000,
         paste0(x / 1000, "k"), x)
}
```

@@

  if (rel) {
    p <- p + 
      geom_hline(yintercept = 1, linetype = "dashed")
  }

@@

# helper function to make subplots E - H
plot_helper <- function(tmp = scores, xvar = x, y, rel = relative, 
                        facet_formula, colour) {
  if (rel) {
    tmp <- tmp %>%
      arrange(target_type, !!sym(xvar)) %>%
      group_by(target_type, !!sym(xvar)) %>%
      mutate_at(y, ~ . / .[Model %in% c("Hub-ensemble", 
                                        "Hub-ensemble-mean")])
  }

  tmp %>%
    ggplot(aes_string(x = xvar, y = y, group = "Model", color = "Model")) +
    geom_point(position = position_dodge(width = 0.1)) +
    geom_line(position = position_dodge(width = 0.1)) +
    facet_wrap(facet_formula, scales = "free_y") +
    theme_minimal() +
    theme(legend.position = "bottom") +
    scale_color_manual(values = colour) + 
    scale_y_continuous(labels = scale_fn) + 
    guides(fill = "none", colour = "none")
}

# function to create Figure 1
aggregate_performance <- function(data, 
                                  x = "horizon",
                                  xlabel = "Forecast horizon",
                                  output_dir = here("analysis", "plots"), 
                                  plotname,
                                  relative = FALSE,
                                  facet_formula = ~ target_type, 
                                  colour_scheme = c("normal", "ensemble"),
                                  vals = NULL) {

@@

  # bias plot in panel F -------------------------------------------------------
  # define appropriate y limits
  bias_limits <- max(abs(scores$bias)) * c(-1.1, 1.1) %>% 
    round(2)
  p_bias <- 
    plot_helper(scores, y = "bias", rel = FALSE, colour = vals, xvar = x,
                facet_formula = facet_formula) + 
    labs(y = "Bias", x = NULL) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    expand_limits(y=bias_limits)

  # plot with absolute error of performance in panel G -------------------------
  p_aem <-
    plot_helper(scores, y = "aem", rel = relative, xvar = x,
                facet_formula = facet_formula, colour = vals) + 
    labs(y = "Absolute error", x = NULL)

  # plot with sd of performance in panel H -------------------------------------
  p_wis_sd <- 
    plot_helper(scores, y = "interval_score_sd", rel = relative, xvar = x,
                facet_formula = facet_formula, colour = vals) + 
    labs(y = "WIS - sd", x = NULL)

@@

  p_cov50 <- scores %>%
    filter(range == c(50)) %>%
    plot_helper(y = "coverage", rel = FALSE, facet_formula = facet_formula, 
                colour = vals, xvar = x) +
    labs(y = "Diff. cov. - 50% PI", x = "") + 
    scale_y_continuous(labels = scales::percent)

@@

  p_cov90 <-  scores %>%
    filter(range == c(90)) %>%
    plot_helper(y = "coverage", rel = FALSE, facet_formula = facet_formula, 
                xvar = x, colour = vals) + 
    labs(y = "Diff. cov. - 90% PI") + 
    scale_y_continuous(labels = scales::percent)

@@

  # put all plots together -----------------------------------------------------
  p_combined4 <- 
    (component_case + component_death) / 
    (p_cov50 + p_cov90) /
    (p_sharpness + p_bias) /
    (p_aem + p_wis_sd) +
    plot_layout(guides = "collect", 
                heights = c(1.3, 1, 1, 1)) &
    plot_annotation(tag_levels = 'A')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin())

@@

# plot version that compares locations for SI
filtered_data[model %in% regular_models & horizon == 2] %>%
  aggregate_performance(plotname = "aggregate-performance-2-weeks-locations-all", 
                        x = "location_name", 
                        xlabel = NULL)

# plot version that compares locations in relative terms for SI
filtered_data[model %in% regular_models & horizon == 2] %>%
  aggregate_performance(plotname = "aggregate-performance-2-weeks-locations-all-rel", 
                        x = "location_name", 
                        xlabel = NULL, relative = TRUE)

# plot for different median ensembles (in relative terms) for SI
filtered_data[model %in% ensemble_models[!grepl("mean", ensemble_models)]] %>%
  # filter(!(model == "Hub-ensemble-with-all" & target_type == "case")) %>%
  aggregate_performance(plotname = "aggregate-performance-rel-ensemble", 
                        relative = TRUE, colour_scheme = "ensemble")

@@

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE)
options(knitr.kable.NA = '')
library(here)
library(magrittr)
library(scoringutils) #needs v0.18
# remotes::install_github("epiforecasts/scoringutils@v0.1.8")
library(knitr)
library(kableExtra)
library(data.table)
library(ggplot2)
library(ggridges)
library(RColorBrewer)
library(patchwork)
library(stringr)
library(covid.german.forecasts)
library(grates)
library(ggdist)
library(ggthemes)
library(scales)
library(dplyr)
library(tidyr)
library(forcats)

@@

# helper function for labeling plots
scale_fn <- function(x) {
  ifelse(abs(x) >= 5000,
         paste0(x / 1000, "k"), x)
}
```

@@

  if (rel) {
    p <- p + 
      geom_hline(yintercept = 1, linetype = "dashed") +
      expand_limits(y = c(0, 1)) +
      scale_y_continuous(breaks = seq(0, max(scores$value), 0.25))
  }

@@

# helper function to make subplots E - H
plot_helper <- function(tmp = scores, xvar = x, y, rel = relative, 
                        facet_formula, colour, scalefree = NULL) {
  if (is.null(scalefree)) {
    scalefree <- "free_y"
  }
  if (rel) {
    scalefree <- "fixed"
    tmp <- tmp %>%
      arrange(target_type, !!sym(xvar)) %>%
      group_by(target_type, !!sym(xvar)) %>%
      mutate_at(y, ~ . / .[Model %in% c("Hub-ensemble", 
                                        "Hub-ensemble-mean")])
  }

  out <- tmp %>%
    ggplot(aes_string(x = xvar, y = y, group = "Model", color = "Model")) +
    geom_point(position = position_dodge(width = 0.1)) +
    geom_line(position = position_dodge(width = 0.1)) +
    theme_minimal() +
    theme(legend.position = "bottom") +
    scale_color_manual(values = colour) + 
    scale_y_continuous(labels = scale_fn) + 
    expand_limits(y = 0) +
    facet_wrap(facet_formula, scales = scalefree) +
    guides(fill = "none", colour = "none")

  return(out)
}

# function to create Figure 1
aggregate_performance <- function(data, 
                                  x = "horizon",
                                  xlabel = "Forecast horizon (weeks)",
                                  output_dir = here("analysis", "plots"), 
                                  plotname,
                                  relative = FALSE,
                                  facet_formula = ~ target_type, 
                                  colour_scheme = c("normal", "ensemble"),
                                  vals = NULL) {

@@

  # bias plot in panel F -------------------------------------------------------
  # define appropriate y limits
  bias_limits <- max(abs(scores$bias)) * c(-1.1, 1.1) %>% 
    round(2)
  p_bias <- 
    plot_helper(scores, y = "bias", rel = FALSE, colour = vals, xvar = x,
                facet_formula = facet_formula, scalefree = "fixed") + 
    labs(y = "Bias", x = NULL) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    expand_limits(y=bias_limits)

  # plot with absolute error of performance in panel G -------------------------
  p_aem <-
    plot_helper(scores, y = "aem", rel = relative, xvar = x,
                facet_formula = facet_formula, colour = vals) + 
    labs(y = "Absolute error", x = xlabel)

  # plot with sd of performance in panel H -------------------------------------
  p_wis_sd <- 
    plot_helper(scores, y = "interval_score_sd", rel = relative, xvar = x,
                facet_formula = facet_formula, colour = vals) + 
    labs(y = "WIS - sd", x = xlabel)

@@

  p_cov50 <- scores %>%
    filter(range == c(50)) %>%
    plot_helper(y = "coverage", rel = FALSE, facet_formula = facet_formula, 
                colour = vals, xvar = x, scalefree = "fixed") +
    labs(y = "Diff. cov. - 50% PI", x = "") + 
    scale_y_continuous(labels = scales::percent)

@@

  p_cov90 <-  scores %>%
    filter(range == c(90)) %>%
    plot_helper(y = "coverage", rel = FALSE, facet_formula = facet_formula, 
                xvar = x, colour = vals, scalefree = "fixed") + 
    labs(y = "Diff. cov. - 90% PI", x = "") + 
    scale_y_continuous(labels = scales::percent)

@@

  # put all plots together -----------------------------------------------------
  p_combined4 <- 
    (component_case + component_death) / 
    (p_cov50 + p_cov90) /
    (p_sharpness + p_bias) /
    (p_aem + p_wis_sd) +
    plot_layout(guides = "collect", 
                heights = c(1.3, 1, 1, 1)) &
    plot_annotation(tag_levels = 'A')  &
    theme(legend.position = 'bottom', 
          legend.box="vertical", legend.margin=margin()) &
    labs(x = xlabel)

@@

# plot version that compares locations for SI
filtered_data[model %in% regular_models & horizon == 2] %>%
  aggregate_performance(plotname = "aggregate-performance-2-weeks-locations-all", 
                        x = "location_name", 
                        xlabel = "Location")

# plot version that compares locations in relative terms for SI
filtered_data[model %in% regular_models & horizon == 2] %>%
  aggregate_performance(plotname = "aggregate-performance-2-weeks-locations-all-rel", 
                        x = "location_name", 
                        xlabel = "Location", relative = TRUE)

# plot for different median ensembles (in relative terms) Figure 4
filtered_data[model %in% ensemble_models[!grepl("mean", ensemble_models)]] %>%
  # filter(!(model == "Hub-ensemble-with-all" & target_type == "case")) %>%
  aggregate_performance(plotname = "aggregate-performance-rel-ensemble", 
                        relative = TRUE, colour_scheme = "ensemble")

@@

# make SI version of Figure 1 for Germany and Poland separately
filtered_data[(model %in% regular_models) & location == "GM" ] %>%
  aggregate_performance(plotname = "aggregate-performance-all-Germany")

filtered_data[(model %in% regular_models) & location == "PL" ] %>%
  aggregate_performance(plotname = "aggregate-performance-all-Poland")

@@

Create comparison of crowd forecasts and baseline model - 2 Week ahead forecasts

dailytruth_data

crowdforecasts <- covid.german.forecasts::prediction_data %>% 
  filter(model == "Crowd forecast")

qs <- c(0.05, 0.25, 0.5, 0.75, 0.95)

basline_forecasts <- truth_data %>%
  mutate(difference = c(NA, diff(log(pmax(true_value, 1)))), 
         target_end_date = as.Date(target_end_date), 
         id = 1:n()) %>%
  rowwise() %>%
  mutate(sd = sd(d$difference[pmax(0, id - (4:1))]), 
         quantile = list(qs)) %>%
  tidyr::unnest(cols = quantile) %>%
  mutate(prediction = exp(qnorm(quantile, mean = log(true_value), sd = sd)), 
         model = "Baseline", 
         forecast_date = target_end_date + 2) %>%
  unique()

plot_df <- prediction_data %>%
  filter(model == "Crowd forecast", 
         quantile %in% qs, 
         grepl("2", target)) %>%
  unique() %>%
  rbind(basline_forecasts, fill = TRUE) %>%
  filter(forecast_date %in% forecast_dates$unfiltered) %>%
  pivot_wider(names_from = quantile, values_from = prediction) %>%
  rename(Model = model) %>%
  mutate(target_type = ifelse(target_type == "case", "Case", "Death"))

plot_df %>%
  ggplot(aes(x = forecast_date, colour = NULL, fill = Model)) +
  geom_line(aes(y = `0.5`)) +
  facet_wrap(location_name ~ target_type, scales = "free_y") +
  geom_ribbon(aes(ymin = `0.05`, ymax = `0.95`), alpha = 0.2) +
  geom_ribbon(aes(ymin = `0.25`, ymax = `0.75`), alpha = 0.5) +
  theme_minimal() +
  scale_y_continuous(labels = scale_fn) +
  labs(y = "Prediction intervals", x = "Forecast date") +
  theme(legend.position = "bottom")

ggsave(here("analysis", "plots", "comparison-forecast-interals.png"), 
       height = 6, width = 9)

Analysis of how data changed over time

```{r}
current_data <- 
  rbind(
    fread(here("data-raw", "daily-incidence-cases.csv")) %>%
      mutate(target_type = "Case"), 
    fread(here("data-raw", "daily-incidence-deaths.csv")) %>%
      mutate(target_type = "Death")
  ) %>%
  filter(location_name %in% c("Germany", "Poland")) %>%
  select(region = location_name, date, value, target_type) %>%
  unique()

original_cases <- original_deaths <- list()
for (target_date in as.character(forecast_dates$unfiltered)) {
  # Get cases  ---------------------------------------------------------------
  original_cases[[target_date]] <- 
    fread(file.path("rt-forecast", "data", "summary", 
                    "cases", target_date, "reported_cases.csv")) %>%
    mutate(fit_date = target_date)

  original_deaths[[target_date]] <- 
    fread(file.path("rt-forecast", "data", "summary", 
                    "deaths", target_date, "reported_cases.csv")) %>%
    mutate(fit_date = target_date)
}
original_cases <- rbindlist(original_cases) %>%
  mutate(target_type = "Case") 
original_deaths <- rbindlist(original_deaths) %>%
  mutate(target_type = "Death")

original <- 
  rbind(original_cases, original_deaths) %>%
  filter(region %in% c("Germany", "Poland")) %>%
  unique() %>%
  rename(value_original = confirm)

dailycomparison <- full_join(current_data, original) %>%
  filter(!is.na(value_original)) %>%
  mutate(difference = value - value_original, 
         rel_difference = difference / value_original)

dailycomparison%>%
  filter(date >= min(forecast_dates$unfiltered -2), 
         date <= max(forecast_dates$unfiltered)) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = difference)) +
  theme_minimal() +
  facet_wrap(region ~ target_type, scales = "free_y") +
  labs(y = "Absolute change in data", x = "Date")

ggsave(here("analysis", "plots", "daily-data-updates.png"), 
       height = 6, width = 9)

# weekly data
weeklycomparison <- dailycomparison %>%
  group_by(date, region, target_type) %>%
  # only take the oldest one for comparison if an update has taken place
  slice_head(n = 1) %>%
  ungroup() %>%
  select(-fit_date) %>%
  unique() %>%
  mutate(week = grates::as_yearweek(as.character(date), firstday = 7)) %>%
  group_by(week, target_type, region) %>%
  summarise(value = sum(value), 
            date = max(date),
            value_original = sum(value_original)) %>%
  mutate(difference = value - value_original, 
         rel_difference = difference / value_original)

weeklycomparison %>%
  filter(date >= min(forecast_dates$unfiltered -2), 
         date <= max(forecast_dates$unfiltered)) %>%
  ggplot(aes(x = date, y = rel_difference)) +
  geom_line() +
  facet_wrap(region ~ target_type, scales = "fixed") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  labs(y = "Relative change in weekly data", x = "Date")

ggsave(here("analysis", "plots", "weekly-data-updates.png"), 
       height = 6, width = 9)

Plot with differences hub ensemble vs. crowd forecasts

```{r}
scores <- eval_forecasts(
  filtered_data[(model %in% regular_models)], 
  summarise_by = c("model", "target_type", "location_name", "horizon", 
                   "forecast_date", "target_phase",
                   "location_target", "classification"),
  compute_relative_skill = FALSE
)

# reformat column
scores[, target_type := str_to_title(target_type)]

p <- scores[model %in% c("Crowd forecast", "Hub-ensemble")] %>%
  filter(horizon == 2) %>%
  select(target_type, model, location_name, forecast_date, interval_score) %>%
  rename(Model = model) %>%
  pivot_wider(names_from = Model, values_from = interval_score) %>%
  ggplot(aes(x = `Crowd forecast` - `Hub-ensemble`)) +
  ggdist::stat_slab(alpha = 1, scale = 1, normalize = "panels") +
  scale_x_continuous(labels = scale_fn) +
  theme(axis.ticks.y = element_blank(), 
        axis.text.y = element_blank(), 
        axis.title.y = element_blank()) +
  scale_y_continuous(labels = NULL) +
  lemon::scale_x_symmetric(labels = scale_fn) +
  theme_minimal() +
  labs(y = "Density", x = "WIS difference Crowd forecast - Hub-ensemble")

p1 <- p +
  facet_wrap(~ target_type, nrow = 2, scales = "free")
p2 <- p +
  facet_wrap(location_name ~ target_type, scales = "free")

p1 + p2

ggsave(here("analysis", "plots", "difference-wis-crowd-ensemble.png"), 
       height = 6, width = 9)

Looking at removed forecasts - yields nothing

files <- list.files(here("crowd-forecast", "raw-forecast-data/"))

raw_forecasts <- purrr::map_dfr(
  files,
  .f = function(x) {
    out <- fread(
      here("crowd-forecast", "raw-forecast-data/", x)
    ) %>%
      mutate(target_end_date = as.character(target_end_date))
    } 
) %>%
  dplyr::filter(location_name %in% c("Germany", "Poland")) %>%
  mutate(location = ifelse(location_name == "Germany", "GM", "PL"), 
         distribution = ifelse(is.na(distribution), "log-normal", distribution), 
         width = ifelse(is.na(width), shape_log_normal, width), 
         target_type = ifelse(is.na(target_type), type, target_type))

# use only the latest forecast from a given forecaster
filtered_forecasts <- raw_forecasts %>%
  group_by(forecaster_id, location, target_type) %>%
  dplyr::filter(forecast_time == max(forecast_time)) %>%
  ungroup()

# replace forecast duration with exact data about forecast date and time
# define function to do this for raw and filtered forecasts
replace_date_and_time <- function(forecasts) {
  forecast_times <- forecasts %>%
    group_by(forecaster_id, location, target_type) %>%
    summarise(forecast_time = unique(forecast_time)) %>%
    ungroup() %>%
    arrange(forecaster_id, forecast_time) %>%
    group_by(forecaster_id) %>%
    mutate(forecast_duration = c(NA, diff(forecast_time))) %>%
    ungroup()

  forecasts <- inner_join(
    forecasts, forecast_times,
    by = c("forecaster_id", "location", "target_type", "forecast_time")) %>%
    mutate(forecast_week = epiweek(forecast_date),
           target_end_date = as.Date(target_end_date)) %>%
    select(-forecast_time)
  return(forecasts)
}

# replace time with duration and date with epiweek
raw_forecasts <- replace_date_and_time(raw_forecasts)
filtered_forecasts <- replace_date_and_time(filtered_forecasts)

# obtain quantiles from forecasts ----------------------------------------------
# define function that returns quantiles depending on condition and distribution
calculate_quantiles <- function(quantile_grid, median, width, forecast_type, 
                                distribution, lower_90, upper_90) {
  if (distribution == "log-normal") {
    values <- list(exp(qnorm(
      quantile_grid, mean = log(as.numeric(median)), sd = as.numeric(width)
      )))
  } else if (distribution == "normal") {
    values <- list((qnorm(
      quantile_grid, mean = (as.numeric(median)), sd = as.numeric(width)
      )))
  } else if (distribution == "cubic-normal") {
    values <- list((qnorm(quantile_grid,
     mean = (as.numeric(median) ^ (1 / 3)), sd = as.numeric(width))
      ) ^ 3)
  } else if (distribution == "fifth-power-normal") {
    values <- list((qnorm(quantile_grid, 
    mean = (as.numeric(median) ^ (1 / 5)), sd = as.numeric(width)
      )) ^ 5)
  } else if (distribution == "seventh-power-normal") {
    values <- list((qnorm(quantile_grid,
     mean = (as.numeric(median) ^ (1 / 7)), sd = as.numeric(width))
     ) ^ 7)
  }
  return(values)
}

quantile_grid <- c(0.01, 0.025, seq(0.05, 0.95, 0.05), 0.975, 0.99)

forecast_quantiles <- filtered_forecasts %>%
  # disregard quantile forecasts this week
  rowwise() %>%
  mutate(quantile = list(quantile_grid),
        value = calculate_quantiles(quantile_grid, median, width, 
        forecast_type, distribution, lower_90, upper_90)) %>%
  unnest(cols = c(quantile, value)) %>%
  ungroup() %>%
  mutate(type = ifelse(target_type == "cases", "case", "death"), 
         target = paste0(horizon, " wk ahead inc ", type), 
         type = "quantile")

# omit forecasters who haven't forecasted at least two targets
forecasters_to_omit <- forecast_quantiles %>%
  select(forecaster_id, location, target_type, forecast_date) %>%
  unique() %>%
  group_by(forecaster_id) %>%
  mutate(n = n(), flag = n >= 2) %>%
  dplyr::filter(!flag) %>%
  pull(forecaster_id) %>%
  unique()
